Blame and Coercion: Together Again for the Second Time

### 2019-10-10 paper.tex
revise intro

In 2007, \citet{herman2010space} proposes a solution to this problem by 
translating Blame Calculus to a variant of \citet{henglein1994dynamic}'s 
Coercion Calculus. Their key idea is to represent casts with coercions, 
which can be composed and normalized. Every coercion looks like a list of 
trees. For normalized coercion, the length of the list is bounded by a small 
constant. And the depths of trees do not grow in composition and 
normalization. These two facts implies that by representing casts with 
coercions, we can restore the space-efficiency. This 
Coercion Calculus, however, does not support blame tracking.

@@

\citet{siek2009exploring} incorporate blame tracking into Coercion Calculus by 
decorating coercions with labels. They also propose that there are four blame 
strategies for their Coercion Calculi:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but also blame more programs.
$ D $ and $ UD $ assign blame labels differently.
Blame strategy is essentially the semantics of casts.
They prove that the \lazyUD{} Coercion Calculus simulates the Blame Calculus 
in \cite{wadler2009well}. But it is unknown whether the simulation also works 
the other way, and what the Blame counterparts of other Coercion Calculi are.

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution is based on Threesome Calculus. 
Their key idea is to represent casts with threesomes, a novel cast 
representation. Threesomes, like coercions, can be composed. There is no 
separate normalization because every threesome is in normal form. 
They prove that their Threesome Calculus bi-simulates
the \lazyUD{} Coercion Calculus and the Blame Calculus, and that threesomes are 
space-efficient.

\citet{siek2012interpretations} present a \lazyD{} Blame Calculus and revised 
lazy Coercion Calculi. They simplify the Coercion Calculi by only working with 
normal coercions (i.e. coercions in normal form). 
Fortunately, coercions are trivially normal when initially constructed, and 
their empowered compose function is conjectured to produces normal coercions as 
well. Also, they conjecture that their Coercion Calculi bi-simulate the 
corresponding Blame Calculi.

Following \citet{siek2010threesomes}, \citet{garcia2013calculating} 
implement all other blame strategies with Threesome Calculus. 
He claims that 
coercion with labels is easy to understand but hard to implement, 
and that 
threesome with labels, however, is easy to implement but hard to understand.
His claim is later affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
The connection 
between his Threesome Calculi and \citet{siek2009exploring}'s Coercion 
Calculi are established by the fact that the former are derived from the latter.

\citet{siek2015blame} revisit the coercion-based approach. They revise the 
\lazyUD{} Coercion Calculus by \citet{siek2012interpretations}, prove its 
space-efficiency and that it bi-simulate the \lazyUD{} Blame Calculus.

Last year, \citet{kuhlenschmidt2018efficient} present Grift, a space-efficient 
and blame-tracking compiler for a gradually typed language of the same name. 
This implementation is based on the \lazyD{} Coercion Calculus by 
\citet{siek2012interpretations}.
Their result suggests that implementing coercion is practical. 
%They implement 
%tuples, which is the first time gradual product types is considered. Their 
%treatment on tuples, however, is shown incorrect 
%\cite{Gradual-TypingType-based-casts}.

@@

So far, it seems that the coercion-based approach is the best way to combine 
space-efficiency and blame-tracking: coercion is claimed easy to 
understand \cite{garcia2013calculating}\cite{siek2015blame} and is shown 
easy enough to implement in a compiler \cite{kuhlenschmidt2018efficient}.
Coercion is not satisfactory, however, in at least three aspects.
Firstly, its compose function is not structurally recursive, which makes 
it difficult to study coercion formally.
For instance, many developers of Grift report that it is tricky to convince 
their proof assistants that composition terminates. 
Secondly, it is desirable to understand the nature of ``normalized cast'', but 
the coercion-based approach capture this concept indirectly because normalized 
coercion is defined by taking a subset of coercion.
Last by not least, the memory representation of normal coercions is not very 
compact.

Perhaps unsurprisingly, threesome has a structurally recursive compose. And its 
definition is self-standing. Together with \citet{garcia2013calculating}'s 
work, they suggest that there can be a cast representation whose definition is 
self-standing, and whose composition is structurally recursive. 
Super-coercion introduced in \citet{garcia2013calculating} could be a promising 
candidate. However, its definition is a bit complicated. It uses 10 
constructors to deal with an elementary type system with only base types and 
function types. And four constructors are directly related to function types. 
Thus, super-coercion might not scale very well to more sophisticated type 
systems.

Now we present one such cast representation: hyper-coercion. As we will show in 
Section \ref{sec:hyper-coercion}, there is a clear connection between it and 
the normalized coercion, so we are optimistic that understanding and 
implementing hyper-coercion should be as easy (or as hard).

@@

We prove formally that the \lazyD{} (resp. \lazyUD{}) hyper-coercion calculus 
bi-simulates the \lazyD{} (resp. \lazyUD{}) Blame Calculus. This is possibly 
the first bi-simulation proof for the \lazyD{} Blame Calculus.

The structure of this paper is as follows. 
Section \ref{sec:blame-calculus} reviews the state-of-art of $Lazy$ Blame 
Calculi. 
In section \ref{sec:hyper-coercion} we present Hyper-coercion.
Section \ref{sec:conclude} concludes.

@@

Blame Calculus is based on Simply Typed Lambda Calculus with sum types and 
product types ($ \mathtt{STLC+} $). 
Let $ S,T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P,Q $ range over pre-types. Every pre-type is a type with a type 
constructor at the top. For simplicity, we only have one base type $ \iota $, 
the unit type. Other pre-types are functions, products, and sums.

@@

%One might be tempted to conclude that inconsistency is the root of all runtime 
%type errors (blames). In the setting of gradual typing, however, runtime 
%type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
top-most type constructor. Shallow-inconsistency is the root of all blames in 
all lazy strategies -- casting a value to a shallowly inconsistent type leads 
to a blame immediately. Shallow-consistency is reflexive, symmetric, but 
not transitive.

@@

\subsection{$Lazy D$ Blame Calculus}

@@

\section{Background: Coercion Calculus}

>>>>

In 2007, \citet{herman2010space} proposes a solution to this problem by 
translating Blame Calculus to a variant of \citet{henglein1994dynamic}'s 
Coercion Calculus. Their key idea is to represent casts with coercions, 
which can be composed and normalized. Every coercion looks like a list of 
trees. For normal coercion, the length of the list is bounded by a small 
constant. And the depths of trees do not grow in composition and 
normalization. These two facts implies that coercion is a space-efficient cast 
representation. This Coercion Calculus, however, does not support blame 
tracking.

@@

\citet{siek2009exploring} incorporate blame tracking into Coercion Calculus by 
decorating coercions with labels. They also propose that there are four blame 
strategies for their Coercion Calculi:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but also blame more programs.
$ D $ and $ UD $ assign blame labels differently.
They prove partially that the \lazyUD{} Coercion Calculus is correct by showing 
that it simulates the Blame Calculus in \cite{wadler2009well}. It is unknown, 
however, whether the simulation also works the other direction.

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution is based on Threesome Calculus. 
Their key idea is to represent casts with threesomes, a novel cast 
representation.
%Threesomes can also be composed. Unlike coercions, there is no 
%separate normalization because every threesome is in normal form. 
They prove that their Threesome Calculus is correct by showing it bi-simulates
the \lazyUD{} Coercion Calculus in \citet{siek2009exploring} and the Blame 
Calculus in \citet{wadler2009well}.
They also prove that threesome is space-efficient as well.

\citet{siek2012interpretations} present the first \lazyD{} Blame Calculus and 
revised lazy Coercion Calculi.
They simplify the Coercion Calculi by only 
working with normal coercions (i.e. coercions in normal form). 
%Fortunately, coercions are trivially normal when initially constructed, and 
%their empowered compose function is conjectured to produces normal coercions 
%as well. 
They conjecture that their Coercion Calculi bi-simulate the corresponding Blame 
Calculi.

Following \citet{siek2010threesomes}, \citet{garcia2013calculating} discovers
Threesome Calculi for other blame strategies. After studying the relation 
between coercion and threesome,
he claims that normal coercion with labels is easy to understand but hard to 
implement and hard to reason about, and that 
threesome with labels, however, is easy to implement and easy to reason about 
but hard to understand. 
His claim is later affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
His threesome calculi are equivalent to the Coercion Calculi in 
\citet{siek2009exploring} because the former ones are derived from the latter 
ones.

\citet{siek2015blame} revisit the coercion-based approach. They revise the 
\lazyUD{} Coercion Calculus in \citet{siek2012interpretations}.
They prove that their calculus agrees with the \lazyUD{} Blame Calculus.

Last year, \citet{kuhlenschmidt2018efficient} present Grift, a compiler for a 
gradually typed language of the same name. 
This implementation is based on the \lazyD{} Coercion Calculus in 
\citet{siek2012interpretations}.
Their result suggests that normal coercion is easy enough to implement in a 
compiler.

@@

So far, it seems that the coercion-based approach is the best way to combine 
space-efficiency and blame-tracking: normal coercion is claimed easy to 
understand \cite{garcia2013calculating}\cite{siek2015blame} and is shown 
easy enough to implement in a compiler \cite{kuhlenschmidt2018efficient}.
Normal coercion is still not satisfactory, however, in at least three aspects.
Firstly, as we mentioned above, it is hard to reason about. The major 
difficulty is from its non-structurally recursive composition.
For instance, many developers of Grift report convincing
their proof assistants that the composition terminates is tricky. 
Secondly, the definition of normal coercion lies unnecessarily on top 
of coercion: all coercions are normal when they are initially constructed, and 
\citet{siek2012interpretations} have shown that there exist functions that 
compose and produce normal coercions.

Perhaps unsurprisingly, threesome has a structurally recursive compose and a 
self-standing definition. Together with \citet{garcia2013calculating}'s 
work, they suggest that there might be a cast representation whose definition 
is self-standing, and whose composition is structurally recursive. 
Super-coercion introduced in \citet{garcia2013calculating} could be a promising 
candidate. However, its definition is a bit complicated. It uses 10 
constructors to deal with an elementary type system with only base types and 
function types. And four constructors are directly related to function types. 
Thus, super-coercion might not scale very well to more sophisticated type 
systems.

We present hyper-coercion, a cast representation whose composition is 
structurally recursive, and whose definition is self-standing. What's more, it 
should be at 
least as easy to implement and understand as normal coercions, because there is 
a clear connection between them, as we will show in Section 
\ref{sec:hyper-coercion}.

By comparing hyper-coercion with normal coercion, we notice another problem of 
the latter in memory representation. \dots

@@

We prove formally that the \lazyD{} (resp. \lazyUD{}) hyper-coercion calculus 
bi-simulates the \lazyD{} (resp. \lazyUD{}) Blame Calculus. This is 
the first bi-simulation proof for the \lazyD{} Blame Calculus as far as we know.

The structure of this paper is as follows. 
Section \ref{sec:blame-calculus} reviews the state-of-art of $Lazy$ Blame 
Calculi. 
Section \ref{sec:coercion-calculus} reviews the state-of-art of $Lazy$ Coercion 
Calculi. 
In section \ref{sec:hyper-coercion} we present Hyper-coercion.
Section \ref{sec:conclude} concludes.

@@

Blame Calculus is based on Simply Typed Lambda Calculus with sum types and 
product types ($ \mathtt{STLC+} $). 
Let $ S,T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P,Q $ range over pre-types. Every pre-type is a type with a type 
constructor at the top. For simplicity, we only have one base type, $ \POOunit 
$. 
Other pre-types are functions, products, and sums.

@@

%One might be tempted to conclude that inconsistency is the root of all runtime 
%type errors (blames). In the setting of gradual typing, however, runtime 
%type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
top-most type constructor. Shallow-inconsistency is the root of all blames in 
all lazy strategies -- casting a value to a shallowly inconsistent type leads 
to a blame. Shallow-consistency is reflexive, symmetric, but 
not transitive.

@@

\subsection{\lazyD{} Blame Calculus}

@@

\section{Background: Coercion Calculus} \label{sec:coercion-calculus}

\todo[inline]{Which version of normal coercion should this section review?}
